---
title: "sobol_with_ode"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)
library(here)

source(here("R", "diffusion.R"))
```

```{r sen}

# come up with first set of sample parameters
# set the number of parameters

np <- 200
initialC = ceiling(runif(np, 10, 20))
nx = ceiling(runif (np, 5, 25))
nt = ceiling(runif (np, 10, 30))
D <- rnorm(mean = 0.5, sd = 0.1, n = np)
area <- rnorm(mean = 10, sd = 5, n = np)
X1 <- cbind.data.frame(initialC = initialC, nx = nx, nt = nt, D = D, area = area)

# repeat to get our second set of samples
np <- 200
initialC = ceiling(runif(np, 10, 20))
nx = ceiling(runif (np, 5, 25))
nt = ceiling(runif (np, 10, 30))
D <- rnorm(mean = 0.5, sd = 0.1, n = np)
area <- rnorm(mean = 10, sd = 5, n = np)
X2 <- cbind.data.frame(initialC=initialC, nx = nx, nt = nt, D = D, area = area)

# fix any negative values as they are not meaningful
X1 <- X1 %>% map_df(pmax, 0.0)
X2 <- X2 %>% map_df(pmax, 0.0)

# create our sobel object and get sets ofparameters for running the model

sens_C <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# our parameter sets are
head(sens_C$X)

# lets add names
colnames(sens_C$X) <- c("initialC", "nx","nt","D", "area")

head(sens_C$X)

parms = as.data.frame(sens_C$X)

result <- diff1(initialC = parms[1,]$initialC, nx = parms[1,]$nx, dx = 1, nt = parms[1,]$nt, dt = 1, D = parms[1,]$D, area = parms[1,]$area)

head(result$conc)

filled.contour(result$conc, xlab = "Time", ylab = "Distance")

# extra our metrics of interest  from this
# max conc it gets to
max(result$conc[nrow(result$conc),])

```
```{r}
# time steps required to get to a threshold concentration
# i want to see when it gets to half of the initial conc.
thresh <- parms[1,]$initialC/2

# i'm just looking at the first segment of the distance, so the first column in the matrix
segment = result$conc[,1]

# which() will tell us all the indices at which this occurs - we will take the first of these
idx <- which( segment <= thresh)[1]

# if it never gets there, i.e. idx is a NULL value, then assign the last time step
idx <- ifelse(is.na(idx), length(result$conc[,1]), idx)

# convert to dataframe and add a time column for ggplot
resdf = as.data.frame(result)
resdf = mutate(resdf, time=seq(1:nrow(resdf)))

ggplot(resdf, aes(x=time, y=conc.1)) +
  geom_line() +
  geom_vline(xintercept = idx, col = "red") +
  labs(y = "Concentration", title = "Time to half initial concentration")

```

# Compute our metric for all the parameter sets

What if we want to run for all parameters

Lets create two additional functions that will help us

* a function that computes the metrics we want

* a function that runs our ode function and computes the metrics (I call it a wrapper function as it is really just a workflow/wrapper to call ode solver and then compute metrics)
```{r sen2}
# turn computing our metrics into a function
# you can add the threshold counter from above, and return idx as well
compute_metrics <- function(result) {
  meanconc <- mean(result$conc[nrow(result$conc),])
  maxconc <- max(result$conc[nrow(result$conc),])
  
  #return list
  list(meanconc = meanconc, maxconc = maxconc)
}

# try it on our first parameter set's result we already have saved
compute_metrics(result)

# now we need to apply the ode and this function for all of our parameters
# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

c_wrapper <- function (initialC, nx, nt, D, area, odefunc, metricfunc, dt=1, dx=1) {
   
    result <- odefunc(initialC = initialC, nx = nx, dx = dx, nt = nt, dt = dt, D = D, area = area)
   
   metrics <- metricfunc(result)
}

# test
print( 
  c_wrapper(initialC=parms[1,]$initialC, nx=parms[1,]$nx, nt=parms[1,]$nt, D=parms[1,]$D, area=parms[1,]$area, odefunc=diff1, metricfunc = compute_metrics)
  )
```



# Next step

Run the wrapper for all parameters and look at results


```{r userwarpper}
# now use pmap as we did before

allresults <- parms %>% pmap(c_wrapper, odefunc = diff1, metricfunc = compute_metrics)

# extract out results from pmap into a data frame
allres <- allresults %>% map_dfr(`[`, c("meanconc", "maxconc"))


# create boxplots
tmp <- allres %>% pivot_longer(cols = everything(), names_to = "metric", values_to = "value")
ggplot(tmp, aes(metric, value, col = metric)) +
  geom_boxplot()
```

# Compute the sobol indicies for each metric

```{r sen3}
#unfortunately there might be NA values, i'm not sure why -- but we need to filter them out
allres = allres %>% filter(!is.nan(maxconc) | !is.nan(meanconc))

# tell() needs a multiple of the number of samples in the original distribution
# in this case, multiple of 200, so we'll need to drop some -- not really ideal
allres = allres[1:1200,]

# sobol can only handle one output at a time  - so we will need to do them separately
sens_C_maxconc <- sensitivity::tell(sens_C, allres$maxconc)

# first-order indices (main effect without co-variance)
rownames(sens_C_maxconc$S) <- names(parms)
sens_C_maxconc$S

# total sensitivity index -note that this partitions the output variance
rownames(sens_C_maxconc$T) <- names(parms)
sens_C_maxconc$T

# create another one for mean concentration
sens_C_meanconc <- sensitivity::tell(sens_C, allres$meanconc)
# first-order indices (main effect without co-variance)
rownames(sens_C_meanconc$S) <- names(parms)
sens_C_meanconc$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
rownames(sens_C_meanconc$T) <- names(parms)
sens_C_meanconc$T
```

