---
title: "Dynamic ODE model"
author: "Haylee Oyler, Jo Cardelle"
format: 
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---

Consider the following model of forest growth (where forest size in measured ni units of carbon (C)) 

• dC/dt = r * C' for forests where C is below a threshold canopy closure
• dC/dt = g* 1( - C/K) for forests where carbon is at or above the threshold canopy closure
• K is a carrying capacity in units of carbon

The size of the forest (C), Canopy closure threshold and carrying capacity are al ni units of carbon You could think of the canopy closure threshold as the size of the forest at which growth rates change from exponential to linear You can think of r, as early exponential growth rate and gas the linear growth rate once canopy closure has been reached

```{r}
library(tidyverse)
library(deSolve)
library(sensitivity)
library(here)

source(here("hw5/dforestgrowth.R"))
source(here("hw5/maxforest.R"))
source(here("hw5/maxforestwrapper.R"))

```


### Forest Growth Model

```{r}
# Initial population
Cinitial <- c(C=10)

# Time period over 300 years
simtimes <- seq(from = 1, to = 300)

# List of parameters
parms <- list(r = 0.01,
              g = 2,
              K = 250,
              thresh = 50)

# Run the ODE model
result <- ode(y = Cinitial, times = simtimes, func = dforestgrowth, parms = parms)
```

## Graphing the Forest Growth Model

```{r}
# Graph forest growth over time
ggplot(result, aes(time, C)) +
  geom_line() +
  geom_hline(yintercept = 50, col = "red") +
  labs(y = "Forest Size (Kg carbon)", 
       title = "Forest Growth Model over 300 Years",
       x = "Years") +
  theme_minimal()
```

## Sobol Sensitivity Analysis

```{r}
# Set initial population 
Cinitial <- 10

# Create parameter distributions
np <- 2000
K <- rnorm(mean = 250, sd = 25, n = np)
r <- rnorm(mean = 0.01, sd = 0.001, n = np)
g <- rnorm(mean = 0.2, sd = 0.02, n = np)
X1 <- cbind.data.frame(r = r, K = K, g = g)
# We don't vary thresh so add it here statically
X1 <- X1 %>% 
  mutate(thresh = 50)
 
# Repeat for second set of samples
np <- 2000
K <- rnorm(mean = 250, sd = 25, n = np)
r <- rnorm(mean = 0.01, sd = 0.001, n = np)
g <- rnorm(mean = 0.2, sd = 0.02, n = np)
X2 <- cbind.data.frame(r = r, K = K, g = g)
# We don't vary thresh so add it here statically
X2 <- X2 %>% 
  mutate(thresh = 50)

# Fix any negative values and they are not meaningful
X1 <- X1 %>% map_df(pmax, 0.0)
X2 <- X2 %>% map_df(pmax, 0.0)

# Run the sobol analysis
sens_C <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# Add parameter names
colnames(sens_C$X) <- c("r", "K", "g", "thresh")
head(sens_C$X)
```

### Use Wrapper Function
The metric we're looking for here is max forest growth. We'll use the functionized version of the `max` and a wrapper function to calculate this value for our model. 

```{r}
# List parameters and format for pmap
parms <- list(r = sens_C$X[, "r"], 
               g = sens_C$X[, "g"],
               K = sens_C$X[, "K"],
               thresh = sens_C$X[, "thresh"])

# Map parameters onto ODE forest growth function
allresults <- parms %>% pmap(maxforestwrapper, Cinitial = c(C = 10), times = simtimes, growth_func = dforestgrowth, max_func = maxforest)

# Extract results and store in a data frame
allres <- allresults %>%
  map_dfr(~ tibble(max_c = .x[[1]]))
```

### Plotting Sobol Results

```{r}
# Retrieve maximum forest size from the sobol output
max_forest <- sens_C$y

# Pivot longer for plotting
allres_long <- allres %>% 
  pivot_longer(cols = everything(), 
               names_to = "max_forest", 
               values_to = "value")

# Visualize results
ggplot(allres_long) + 
  geom_boxplot(aes(x = max_forest, y = value)) +
  labs(title = "Sobol Analysis of Maximum Forest Size",
       y = "Maximum Forest Size (Kg C)",
       x = "") +
  theme_minimal()
```
### Sobel Indices
```{r}
# Get indices for max forest
sens_Cmax <- sensitivity::tell(sens_C, allres$max_c)

# First-order indices 
rownames(sens_Cmax$S) <- c("r", "g", "K", "thresh")
print(sens_Cmax$S)

# Total sensitivity index
rownames(sens_Cmax$T) <- c("r", "g", "K", "thresh")
print(sens_Cmax$T)
```

### Conclusions

From the results of our Sobol analysis, the model is most sensitive to the growth rate r and next most sensitive to the carrying capacity K. It has no significant sensitivity to thresh, which makes sense because we did not vary it. It also has a small sensitivity to the growth rate g. We see similar results across both the first order indices and the total sensitivity index, although they are orders of magnitude apart from each other. Climate change could have a large impact on the most sensitive parameters. For example, higher amounts of CO2 in the atmosphere could increase the growth rate r, or drier overall conditions could affect what the carrying capacity in a locale is.
